Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-07 Final_Draft.Rmd
rp coefficient plot work and formatting

```{r fs_densityplot, echo = FALSE}

@@

In the second regression, the results are similar. The main coefficient on segregation index has decreased in magnitude, although in these specifications, it remains statistically significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Latino, and median household income have all switched signs.

```{r results_printing_rp_impute_regressions_2, results = "asis", echo=FALSE}

@@

In the third regression, like in the second, the main coefficient on segregation index has decreased in magnitude but remains significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Asian, median income, and logged population have switched signs. In all three of the specifications, tandard errors reduced slightly using the larger, multiply imputed datasets compared to the original analysis with the original data in @trounstine_segregation_2016.

```{r results_printing_rp_impute_regressions_3, results = "asis", echo=FALSE}

@@

@@

The results of the first regression present a stark difference to the results of @trounstine_segregation_2016. Crucially, the effect size of the segregation index on public spending has essentially gone to zero and has become statistically insignificant. Interestingly, the coefficient on the diversity variable has increased substantially from .106 to .759 and has become statistically significant. The sign of the coefficient on percent Asian popultion has switched signs from negative to positive; however, it remains statistically insignificant.

@@

```{r results_printing_fs_impute_regressions_2, results = "asis", echo=FALSE, eval = FALSE, invlude = FALSE}

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo=FALSE, eval = FALSE, include = FALSE}

@@

@@

```{r dataset_rows_in_analysis}

@@

```{r dataset_medians}

@@

model_summaries %>%
  ggplot(aes(x = median, y = estimate)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable, labeller = labeller(variable = coef.labels)) +
  ggtitle("Coefficient Plots for Diversity and White/Non-White Segregation Indices Across\nthe Models")+
  xlab("Models Ordered by Increasing Median Population Size of the Sample") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public\ngoods spending varies with city size. As city populaiton size increases, diversity tends to have\nless explanatory power for public goods spending while segregation has more. The opposite\nis true when considering city population size decreasing: segregation becomes less important\nas diversity becomes increasingly important.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0))

@@

Why cities differ in their amount of spending on public goods is an interesting questions that is frequently debated in the literature. @trounstine_segregation_2016 posits that segregation may be an important underlying cause of the underprovsion of public goods. Finding evidence that residential segregation by race increases political partisanship, @trounstine_segregation_2016 argues that residential segregation can increase division and make cooperation amongst competing groups difficult, which results in decreased spending on public goods. The findings of @trounstine_segregation_2016 are particularly important because they suggest that segregation, as defined by many homogeneous neighborhoods within a larger diverse geographical area, is the key factor in disinvestment in public goods. This contradicts previous research which has suggested that levels of diversity were instead most important (see for example @baqir_public_1999 or @hopkins_diversity_2009). While @trounstine_segregation_2016 still represents relatively new reasearch as of 2020, it has yet to be fully vetted by additional research. Furthermore, some research that has been published on the topic since its publication suggests that other factors such as larger inequities in the political system favoring socially powerful groups [@lee_ethnic_2018] or the interaction of income and racial inequality [@an_its_2018] may be more important in explaining public goods provisions.

@@

```{r  include = FALSE}

@@

```{r felm, include = FALSE}

@@

```{r main_fp_tables, results = "asis", echo = FALSE}

@@

>>>>

@@

In the second regression, the results are similar. The main coefficient on segregation index has decreased in magnitude, it remains statistically significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Latino, and median household income have all switched signs.

```{r results_printing_rp_impute_regressions_2, results = "asis", include = FALSE, eval = FALSE}

@@

In the third regression, like in the second, the main coefficient on segregation index has decreased in magnitude but remains significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Asian, median income, and logged population have switched signs. In all three of the specifications, standard errors reduced slightly using the larger, multiply imputed datasets compared to the original analysis with the original data in @trounstine_segregation_2016.

```{r results_printing_rp_impute_regressions_3, results = "asis", include = FALSE, eval = FALSE}

@@

```{r Trounstine_rp_models_for_coeff_plot, include = FALSE}

m1_estimate_seg <- 0.931958
m1_se_seg <- 0.394180
m1_estimate_d <- 0.384535
m1_se_d <-0.361625

m2_estimate_seg <- 0.755719
m2_se_seg <- 0.296916 
m2_est_d <- 0.518369
m2_se_d <- 0.323161

m3_estimate_seg <- 0.835441
m3_se_seg <- 0.295759
m3_est_d <- 0.583684
m3_se_d <- 0.323488

```{r calc_coeff_sds_original_rp_impute, include = FALSE}

model_coeff_rp <- tibble(
  variable = c("Segregation Index", "Diversity", 
               "Segregation Index", "Diversity",
               "Segregation Index", "Diversity", 
               "Segregation Index", "Diversity",
                "Segregation Index", "Diversity", 
               "Segregation Index", "Diversity"),
  estimate = c(imp_1_rp_sum[2,2], imp_1_rp_sum[3,2],
               imp_2_rp_sum[2,2], imp_2_rp_sum[3,2],
               imp_3_rp_sum[2,2], imp_3_rp_sum[3,2],
               m1_estimate_seg, m1_estimate_d,
               m2_estimate_seg, m2_est_d,
               m3_estimate_seg, m3_est_d),
  std_error = c(imp_1_rp_sum[2,3], imp_1_rp_sum[3,3],
                imp_2_rp_sum[2,3], imp_2_rp_sum[3,3],
                imp_3_rp_sum[2,2], imp_3_rp_sum[3,2],
                m1_se_seg, m1_se_d,
                m2_se_seg, m2_se_d,
                m3_se_seg, m3_se_d),
  model = c( "Imputation 1", "Imputation 1",
             "Imputation 2","Imputation 2",
             "Imputation 3", "Imputation 3",
             "Trounstine 1", "Trounstine 1",
            "Trounstine 2", "Trounstine 2",
            "Trounstine 3",  "Trounstine 3"),
  order = c (2,2,
             5,5,
             8,8,
             1,1,
             4,4,
             7,7),
  color = c("Imputations", "Imputations", "Imputations", 
            "Imputations", "Imputations", "Imputations", 
            "Trounstine", "Trounstine", "Trounstine",
            "Trounstine","Trounstine","Trounstine"),
  model_num = c("1","1","2","2","3","3","1","1","2","2","3","3")
)

model_coeff_rp <- model_coeff_rp  %>%
  mutate(order = as.factor(order))

model_coeff_rp %>%
  ggplot(aes(x = order, y = estimate, color = color, shape = model_num)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable) +
  ggtitle("Coefficient Plots for Diversity and Segregation Indices Across\nthe Racial Polarization Models")+
  xlab("Model Specifications") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public goods spending varies\n with city size. As city populaiton size increases, diversity tends to have less explanatory power for public goods\nspending while segregation has more. The opposite is true when considering city population size decreasing:\nsegregation becomes less important as diversity becomes increasingly important.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0),
        axis.ticks.x = element_blank()) +
  scale_x_discrete(labels = c("            1", "", "            2", "", "           3", "")) +
  labs(color = "Model Category", shape = "Specification") +
  scale_color_manual(values= c("orange", "blue"))

@@

The results of the first regression present a stark difference to the results of @trounstine_segregation_2016. Crucially, the effect size of the segregation index on public spending has essentially gone to zero and has become statistically insignificant. Interestingly, the coefficient on the diversity variable has increased substantially from .106 to .759 and has become statistically significant. The results of the second and third specifications were similar, with the coefficient on segregation going almost to zero and becoming insignificant on both. For brevity, these results are not presented.

@@

```{r results_printing_fs_impute_regressions_2, results = "asis", echo=FALSE, eval = FALSE, include = FALSE}

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo = FALSE, eval = FALSE, include = FALSE}

@@

```{r run_first_T_mode_for_coeff_plot, include = F}
# Main Analysis 2

felm1 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)
```

@@

```{r dataset_rows_in_analysis, include = FALSE}

@@

```{r dataset_medians, include = FALSE}

@@

model_summaries %>%
  ggplot(aes(x = median, y = estimate)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable) +
  ggtitle("Coefficient Plots for Diversity and White/Non-White Segregation Indices Across\nthe Models")+
  xlab("Models Ordered by Increasing Median Population Size of the Sample") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public goods spending varies\n with city size. As city populaiton size increases, diversity tends to have less explanatory power\nfor public goods spending while segregation has more. The opposite is true when considering\ncity population size decreasing: segregation becomes less important as diversity becomes\nincreasingly important.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0))

@@

Why cities differ in their amount of spending on public goods is an interesting question that is frequently debated in the literature. @trounstine_segregation_2016 posits that segregation may be an important underlying cause of the underprovsion of public goods. Finding evidence that residential segregation by race increases political partisanship, @trounstine_segregation_2016 argues that residential segregation can increase division and make cooperation amongst competing groups difficult, which results in decreased spending on public goods. The findings of @trounstine_segregation_2016 are particularly important because they suggest that segregation, as defined by many homogeneous neighborhoods within a larger diverse geographical area, is the key factor in disinvestment in public goods. This contradicts previous research which has suggested that levels of diversity were instead most important (see for example @baqir_public_1999 or @hopkins_diversity_2009). While @trounstine_segregation_2016 still represents relatively new reasearch as of 2020, it has yet to be fully vetted by additional research. Furthermore, some research that has been published on the topic since its publication suggests that other factors such as larger inequities in the political system favoring socially powerful groups [@lee_ethnic_2018] or the interaction of income and racial inequality [@an_its_2018] may be more important in explaining public goods provisions.

@@

@@

```{r felm, include = FALSE, eval = FALSE}

@@

```{r main_fp_tables, results = "asis", echo = FALSE, eval = FALSE, include = FALSE}

@@

```{r fs_densityplot, echo = FALSE, eval = FALSE, include = FALSE}
